---
title: "Correlation of RNP-MaP Runs"
format: html
editor: visual
---

## Correlation of RNP-MaP Runs

We are comparing the SHAPE outputs of the mutation rate to untreated rates and looking at the correlation between the two replicates.

Loading in Libraries

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/jarrettscomputer/Desktop/RNP_MaP_Data/Shape_MaP_Output_profiles")
library(ggplot2)
library(dplyr)
```

Data

```{r}
#U1
U1_Rep1_aq_int <- read.table('U1_aq_int_Hs_RNU1_1_profile.txt', header = TRUE)
U1_Rep2_aq_int <- read.table('U1_aq_int_20250707_Hs_RNU1_1_profile.txt', header = TRUE)
U1_Rep1_DMSO_int<- read.table('U1_DMSO_int_Hs_RNU1_1_profile.txt', header = TRUE)
U1_Rep2_DMSO_int <- read.table('U1_DMSO_int_20250707_Hs_RNU1_1_profile.txt', header = TRUE)
#U2
U2_Rep2_aq_int <- read.table('U2_aq_int_20250707_Hs_RNU2_1_profile.txt', header = TRUE)
U2_Rep1_aq_int<- read.table('U2_aq_int_Hs_RNU2_1_profile.txt', header = TRUE)
U2_Rep2_DMSO_int<- read.table('U2_DMSO_int_20250707_Hs_RNU2_1_profile.txt', header = TRUE)
U2_Rep1_DMSO_int<- read.table('U2_DMSO_int_Hs_RNU2_1_profile.txt', header = TRUE)
#U11
U11_Rep2_aq_int<- read.table('U11_aq_int_20250707_Hs_RNU11_profile.txt', header = TRUE)
U11_Rep1_aq_int<- read.table('U11_aq_int_Hs_RNU11_profile.txt', header = TRUE)
U11_Rep2_DMSO_int<- read.table('U11_DMSO_int_20250707_Hs_RNU11_profile.txt', header = TRUE)
U11_Rep1_DMSO_int<- read.table('U11_DMSO_int_Hs_RNU11_profile.txt', header = TRUE)
#U12
U12_Rep1_aq_int<- read.table('U12_aq_int_20250707_Hs_RNU12_profile.txt', header = TRUE)
U12_Rep2_aq_int<- read.table('U12_aq_int_Hs_RNU12_profile.txt', header = TRUE)
U12_Rep1_DMSO_int<- read.table('U12_DMSO_int_20250707_Hs_RNU12_profile.txt', header = TRUE)
U12_Rep2_DMSO_int<- read.table('U12_DMSO_int_Hs_RNU12_profile.txt', header = TRUE)

```

Filtering Data & Defining Columns

```{r}
#U1_aq_int
U1_Rep1_subset <- U1_Rep1[, c("Modified_rate", "Untreated_rate")]
U1_Rep2_subset <- U1_Rep2[, c("Modified_rate", "Untreated_rate")]

U1_Rep1_subset <- U1_Rep1_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>%
  mutate(across(c(Ratio, Log_10_Ratio), ~ ifelse(is.infinite(.), NA, .)))

Combined_ratios_U1 <- data.frame(
  Rep1 = U1_Rep1_subset$Log_10_Ratio,
  Rep2 = U1_Rep2_subset$Log_10_Ratio
)

Cleaned_Combined_ratios_U1 <- na.omit(Combined_ratios_U1)

#U1_DMSO_int
U1_Rep1_subset <- U1_Rep1[, c("Modified_rate", "Untreated_rate")]
U1_Rep2_subset <- U1_Rep2[, c("Modified_rate", "Untreated_rate")]

U1_Rep1_subset <- U1_Rep1_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>%
  mutate(across(c(Ratio, Log_10_Ratio), ~ ifelse(is.infinite(.), NA, .)))

Combined_ratios_U1 <- data.frame(
  Rep1 = U1_Rep1_subset$Log_10_Ratio,
  Rep2 = U1_Rep2_subset$Log_10_Ratio
)

Cleaned_Combined_ratios_U1 <- na.omit(Combined_ratios_U1)

#U2_aq_int
U1_Rep1_subset <- U1_Rep1[, c("Modified_rate", "Untreated_rate")]
U1_Rep2_subset <- U1_Rep2[, c("Modified_rate", "Untreated_rate")]

U1_Rep1_subset <- U1_Rep1_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>%
  mutate(across(c(Ratio, Log_10_Ratio), ~ ifelse(is.infinite(.), NA, .)))

Combined_ratios_U1 <- data.frame(
  Rep1 = U1_Rep1_subset$Log_10_Ratio,
  Rep2 = U1_Rep2_subset$Log_10_Ratio
)

Cleaned_Combined_ratios_U1 <- na.omit(Combined_ratios_U1)
#U2_DMSO_int
U1_Rep1_subset <- U1_Rep1[, c("Modified_rate", "Untreated_rate")]
U1_Rep2_subset <- U1_Rep2[, c("Modified_rate", "Untreated_rate")]

U1_Rep1_subset <- U1_Rep1_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>%
  mutate(across(c(Ratio, Log_10_Ratio), ~ ifelse(is.infinite(.), NA, .)))

Combined_ratios_U1 <- data.frame(
  Rep1 = U1_Rep1_subset$Log_10_Ratio,
  Rep2 = U1_Rep2_subset$Log_10_Ratio
)

Cleaned_Combined_ratios_U1 <- na.omit(Combined_ratios_U1)
#U11_aq_int
U1_Rep1_subset <- U1_Rep1[, c("Modified_rate", "Untreated_rate")]
U1_Rep2_subset <- U1_Rep2[, c("Modified_rate", "Untreated_rate")]

U1_Rep1_subset <- U1_Rep1_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>%
  mutate(across(c(Ratio, Log_10_Ratio), ~ ifelse(is.infinite(.), NA, .)))

Combined_ratios_U1 <- data.frame(
  Rep1 = U1_Rep1_subset$Log_10_Ratio,
  Rep2 = U1_Rep2_subset$Log_10_Ratio
)

Cleaned_Combined_ratios_U1 <- na.omit(Combined_ratios_U1)
#U11_DMSO_int
U1_Rep1_subset <- U1_Rep1[, c("Modified_rate", "Untreated_rate")]
U1_Rep2_subset <- U1_Rep2[, c("Modified_rate", "Untreated_rate")]

U1_Rep1_subset <- U1_Rep1_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>%
  mutate(across(c(Ratio, Log_10_Ratio), ~ ifelse(is.infinite(.), NA, .)))

Combined_ratios_U1 <- data.frame(
  Rep1 = U1_Rep1_subset$Log_10_Ratio,
  Rep2 = U1_Rep2_subset$Log_10_Ratio
)

Cleaned_Combined_ratios_U1 <- na.omit(Combined_ratios_U1)
#U12_aq_int
U1_Rep1_subset <- U1_Rep1[, c("Modified_rate", "Untreated_rate")]
U1_Rep2_subset <- U1_Rep2[, c("Modified_rate", "Untreated_rate")]

U1_Rep1_subset <- U1_Rep1_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>%
  mutate(across(c(Ratio, Log_10_Ratio), ~ ifelse(is.infinite(.), NA, .)))

Combined_ratios_U1 <- data.frame(
  Rep1 = U1_Rep1_subset$Log_10_Ratio,
  Rep2 = U1_Rep2_subset$Log_10_Ratio
)

Cleaned_Combined_ratios_U1 <- na.omit(Combined_ratios_U1)
#U12_DMSO_int
U1_Rep1_subset <- U1_Rep1[, c("Modified_rate", "Untreated_rate")]
U1_Rep2_subset <- U1_Rep2[, c("Modified_rate", "Untreated_rate")]

U1_Rep1_subset <- U1_Rep1_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>% mutate(Ratio = Modified_rate / Untreated_rate) %>% mutate(Log_10_Ratio = log10(Modified_rate / Untreated_rate)) %>%
  mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))

U1_Rep2_subset <- U1_Rep2_subset %>%
  mutate(across(c(Ratio, Log_10_Ratio), ~ ifelse(is.infinite(.), NA, .)))

Combined_ratios_U1 <- data.frame(
  Rep1 = U1_Rep1_subset$Log_10_Ratio,
  Rep2 = U1_Rep2_subset$Log_10_Ratio
)

Cleaned_Combined_ratios_U1 <- na.omit(Combined_ratios_U1)
```

```{r}
process_replicates <- function(rep1, rep2) {
  rep1_subset <- rep1[, c("Modified_rate", "Untreated_rate")] %>%
    mutate(Ratio = Modified_rate / Untreated_rate,
           Log_10_Ratio = log10(Ratio)) %>%
    mutate(across(everything(), ~ ifelse(. == 0, NaN, .)))%>%
    mutate(across(c(Ratio, Log_10_Ratio), ~ ifelse(is.infinite(.), NA, .)))

  rep2_subset <- rep2[, c("Modified_rate", "Untreated_rate")] %>%
    mutate(Ratio = Modified_rate / Untreated_rate,
           Log_10_Ratio = log10(Ratio)) %>%
    mutate(across(everything(), ~ ifelse(. == 0, NaN, .))) %>%
    mutate(across(c(Ratio, Log_10_Ratio), ~ ifelse(is.infinite(.), NA, .)))
  
  combined <- data.frame(
    Rep1 = rep1_subset$Log_10_Ratio,
    Rep2 = rep2_subset$Log_10_Ratio
  )

  cleaned <- na.omit(combined)
  return(cleaned)
}

#Change sample names here!!!!!!!!!
samples <- c("U1", "U2", "U11", "U12")
conditions <- c("aq", "DMSO")

results <- list()

for (sample in samples) {
  for (cond in conditions) {
    rep1_name <- paste0(sample, "_Rep1_", cond, "_int")
    rep2_name <- paste0(sample, "_Rep2_", cond, "_int")

    rep1 <- get(rep1_name)
    rep2 <- get(rep2_name)

    cleaned_data <- process_replicates(rep1, rep2)

    key <- paste0(sample, "_", cond)
    results[[key]] <- cleaned_data
  }
}



```

Testing correlation and Analyzing the Linear Relationship (further from zero the stronger the relationship)

```{r}
#U1
shapiro.test(Combined_ratios_U1$Rep1)
shapiro.test(Combined_ratios_U1$Rep2)
cor.test(Combined_ratios_U1$Rep1,Combined_ratios_U1$Rep2 , method = "pearson")
cov(x = Cleaned_Combined_ratios_U1$Rep1, y = Cleaned_Combined_ratios_U1$Rep2)

#U2
shapiro.test(Combined_ratios_U1$Rep1)
shapiro.test(Combined_ratios_U1$Rep2)
cor.test(Combined_ratios_U1$Rep1,Combined_ratios_U1$Rep2 , method = "pearson")
cov(x = Cleaned_Combined_ratios_U1$Rep1, y = Cleaned_Combined_ratios_U1$Rep2)

#U11
shapiro.test(Combined_ratios_U1$Rep1)
shapiro.test(Combined_ratios_U1$Rep2)
cor.test(Combined_ratios_U1$Rep1,Combined_ratios_U1$Rep2 , method = "pearson")
cov(x = Cleaned_Combined_ratios_U1$Rep1, y = Cleaned_Combined_ratios_U1$Rep2)

#U12
shapiro.test(Combined_ratios_U1$Rep1)
shapiro.test(Combined_ratios_U1$Rep2)
cor.test(Combined_ratios_U1$Rep1,Combined_ratios_U1$Rep2 , method = "pearson")
cov(x = Cleaned_Combined_ratios_U1$Rep1, y = Cleaned_Combined_ratios_U1$Rep2)
```

Spearman Test

```{r}
run_all_tests <- function(df, sample_name) {

  spearman_result <- cor.test(df$Rep1, df$Rep2, method = "spearman")
  spearman_rho <- spearman_result$estimate
  
  list(
    Shapiro_Rep1_p = shapiro.test(df$Rep1)$p.value,
    Shapiro_Rep2_p = shapiro.test(df$Rep2)$p.value,
    spearman_value = spearman_rho,
    Covariance = cov(df$Rep1, df$Rep2)
  )
}

test_results_spearman <- lapply(results, run_all_tests)

test_summary_spearman  <- do.call(rbind, test_results_spearman)
test_summary_spearman  <- as.data.frame(test_summary_spearman)
test_summary_spearman $Sample <- rownames(test_summary_spearman)
rownames(test_summary_spearman ) <- NULL
#print(test_summary)


```

Pearson Test

```{r}
run_all_tests <- function(df, sample_name) {
  qqnorm(df$Rep1, main = paste("Q-Q Plot for", sample_name, "Rep1"))
  qqline(df$Rep1, col = "green")

  qqnorm(df$Rep2, main = paste("Q-Q Plot for", sample_name, "Rep2"))
  qqline(df$Rep2, col = "blue")

  pearson_result <- cor.test(df$Rep1, df$Rep2, method = "pearson")
  pearson_r <- pearson_result$estimate

  list(
    Shapiro_Rep1_p = shapiro.test(df$Rep1)$p.value,
    Shapiro_Rep2_p = shapiro.test(df$Rep2)$p.value,
    pearson_value = pearson_r,
    R_Squared_Value = pearson_r^2,
    Covariance = cov(df$Rep1, df$Rep2)
  )
}

sample_names <- names(results)
test_results_pearson <- mapply(run_all_tests, results, sample_names, SIMPLIFY = FALSE)
test_summary_pearson <- do.call(rbind, test_results_pearson)
test_summary_pearson <- as.data.frame(test_summary_pearson)
test_summary_pearson$Sample <- sample_names
#print(test_summary)
```

## Observing the correlation plots of mod_rates (mod/unt_rates)

Super_plot_non_linear

```{r}
for (name in names(results)) {
  df <- results[[name]]

  Above_ground_pool <- ggplot(df, aes(x = Rep1, y = Rep2)) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE, color = "blue") +
    labs(title = name, x = "Rep_1 (Log10 Ratio)", y = "Rep_2 (Log10 Ratio)") +
    theme_minimal()

  print(Above_ground_pool)
}
```

Super_Plot_Linear

```{r}

for (name in names(results)) {
  df <- results[[name]]

  Above_ground_pool <- ggplot(df, aes(x = Rep1, y = Rep2)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, color = "blue") +
    labs(title = name, x = "Rep_1 (Log10 Ratio)", y = "Rep_2 (Log10 Ratio)") +
    theme_minimal()

  print(Above_ground_pool)
}
```

U1

```{r}
ggplot(Cleaned_Combined_ratios_U1, aes(x = Cleaned_Combined_ratios_U1$Rep1, y =Cleaned_Combined_ratios_U1$Rep2)) + geom_point() + geom_smooth(method = "lm", se = FALSE, color = "blue") 
```

U2

```{r}
ggplot(Cleaned_Combined_ratios_U1, aes(x = Cleaned_Combined_ratios_U1$Rep1, y =Cleaned_Combined_ratios_U1$Rep2)) + geom_point() + geom_smooth(method = "lm", se = FALSE, color = "blue") 
```

U11

```{r}
ggplot(Cleaned_Combined_ratios_U1, aes(x = Cleaned_Combined_ratios_U1$Rep1, y =Cleaned_Combined_ratios_U1$Rep2)) + geom_point() + geom_smooth(method = "lm", se = FALSE, color = "blue") 
```

U12

```{r}
ggplot(Cleaned_Combined_ratios_U1, aes(x = Cleaned_Combined_ratios_U1$Rep1, y =Cleaned_Combined_ratios_U1$Rep2)) + geom_point() + geom_smooth(method = "lm", se = FALSE, color = "blue") 
```

Answer: Modfied rate is looking similar from both samples as indicated by the examples

What are the features of the site that makes it a good SM Binder

K_mean clustering

Dynamic Time Warping??

We know what a protein binding site on a rna looks like, but we know that the signal will not change even if there is an insert. He said never mind, back to this later.
